---
title: "Lesson1"
author: "mom"
date: "5/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# ALT + COMMAND + i gives you a new code chunk:
```{r second}
#testing
x <- c(1, 2, 3)
```
The mean of x is `r mean(x)`. <i>THIS IS ITALICS</i>.

Here is a handy reference for tables:
https://bookdown.org/yihui/rmarkdown-cookbook/kable.html

[Here is a cheatsheet for R Markdown](https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)

# How to push changes to GitHub:
Note to self: This only works if you have cloned the repository to your computer and are pushing to the same repository on GitHub. The advantage is that you can create code in the R Studio environment and then push to GitHub.
1) Save .Rmd file 
From Terminal:
2) cd /Users/lja001/Documents/Test
3) git add Lesson1.Rmd
4) git commit -m 'add R Markdown'
5) git push

# Presentation from Athanasia Mowinckel:
[The slides and source code can be found here:](https://github.com/rladies/meetup-presentations_oslo/tree/master/2020-05-11-rmarkdown_like_a_boss)
